Parameter estimation

### 2023-05-03 SFE-article/Sections/Experiments.tex
03.05

	The caraway seeds used in the experiments were obtained from Caraway Finland during the 2022 harvesting season. The experiments were conducted using a 10-litre extractor with an inner diameter of approximately 15 cm and a height of 60 cm, which was rented from VTT Bioruukki. Four experiments were performed under different operating conditions: $40^\circ C / 200$ bar, $50^\circ C / 200$ bar, $40^\circ C / 300$ bar, and $50^\circ C / 300$ bar. The volumetric flow rate used in all experiments was 0.4 litres per minute, with negligible variations. The amount of solid material used for extraction was 1 kg (or 1.6 litres), which was not enough to fill the entire extraction chamber.

@@

	The results of experiments are presented in the table \ref{tab: Yield_data}.

@@

		\begin{tabular}{c|cccc}

@@

			150      & 66.8                    & 67.7                    & 74.6                    & 74.9                   

>>>>

	The caraway seeds used in the experiments were obtained from Caraway Finland during the 2022 harvesting season. The experiments were conducted using a 10-litre extractor with an inner diameter of approximately 15 cm and a height of 60 cm, which was rented from VTT Bioruukki. Four experiments were performed under different operating conditions: $40^\circ C / 200$ bar, $50^\circ C / 200$ bar, $40^\circ C / 300$ bar, and $50^\circ C / 300$ bar. The volumetric flow rate used in all experiments was on average 0.4 litres per minute, with negligible variations (up to 5\%). The amount of solid material used for extraction was 1 kg (or 1.6 litres), which was not enough to fill the entire extraction chamber.

@@

	The results of experiments are presented in the Table \ref{tab: Yield_data}.

@@

		\begin{tabular}{ccccc}
			\hline

@@

			150      & 66.8                    & 67.7                    & 74.6                    & 74.9              		 \\ \hline     

### 2023-05-03 SFE-article/Sections/Model.tex
03.05

		\subsubsection{Extraction yield} \label{CH: Yield} \todo{Give reference for measurement function}

		The efficiency of the process (the yield) is calculated according to equations \ref{Model_measurment_1} to \ref{Model_measurment_2}, which evaluate the mass of solute at the exit of the extraction unit and sums it. The integral form of the measurement equation can be transformed into the differential form and augmented with model equations to be solved simultaneously.

@@

			In a state-space sense, the state variables of the system are the local concentrations of solute in the fluid and solid phases (${\color{blue}c_f}(t,z)$ and ${\color{blue}c_s}(t,z)$, respectively), and the local enthalpy of the pseudo-homogeneous phase (${\color{blue}h}(t,z)$). The controllable input variables are the mass flow-rate and temperature of the solvent in the feed ($F_\text{in}(t) = {\color{red}P}(t)$ and $h_\text{in}(t) = h(t,z=0)$, respectively) and the pressure in the extractor ($P(t,z) ={\color{red}P}(t)$). We extend the system state-space by assuming that extraction yield can be modelled as a function of a known initial mass of solute in the solid phase and it can be measured after the separator ($Y(t)$). The system is controllable by manipulating the flow-rate and temperature (enthalpy) of $CO_2$ in the feed, and the pressure in the extractor.

>>>>

		\subsubsection{Extraction yield} \label{CH: Yield}

		The efficiency of the process (the yield) is calculated according to equations \ref{Model_measurment_1} as presented by \citet{Sovova1994a}. The measurement equation evaluate the mass of solute at the outlet of the extraction unit and sums it. The integral form of the measurement equation (\ref{Model_measurment_1}) can be transformed into the differential form (\ref{Model_measurment_2}) and augmented with the process model.

@@

			In a state-space sense, the state variables of the system are the local concentrations of solute in the fluid and solid phases (${\color{blue}c_f}(t,z)$ and ${\color{blue}c_s}(t,z)$, respectively), and the local enthalpy of the pseudo-homogeneous phase (${\color{blue}h}(t,z)$). The controllable input variables are the mass flow-rate and temperature of the solvent in the feed ($F_\text{in}(t) = {\color{red}P}(t)$ and $h_\text{in}(t) = {\color{red}h}(t,z=0)$, respectively) and the pressure in the extractor ($P(t,z) ={\color{red}P}(t)$). We extend the system state-space by assuming that extraction yield can be modelled as a function of a known initial mass of solute in the solid phase and it can be measured after the separator ($Y(t)$). The system is controllable by manipulating the flow-rate and temperature (enthalpy) of $CO_2$ in the feed, and the pressure in the extractor.

### 2023-05-03 SFE-article/Sections/Results.tex
03.05

    The parameter estimation problem was solved by fitting the process model to the dataset given in the table \ref{tab: Yield_data}. Each time-series was fitted to the model separately. As a result of fitting, the following parameters were obtained:

@@

	To ensure that parameters found by the optimizer do not reach unrealistic values, an additional set of inequality constraints is introduced. The lower and upper bounds for each parameter are given in Table \ref{tab:Constraints}.

	\begin{table}[!h]
		\centering
		\adjustbox{max width=\columnwidth}{%
		\begin{tabular}{ lcccccc }
			\hline 
			Parameter	&$k_m$ 	& $D_i$ 	& $D_e^M$ 	& $m_{total}$	& $\tau$ 	& $\sigma$ \\  \hline
			Lower bound	&0	  	& 0 	  	& 0 		& 76 		 	& 0 	   	& 0 \\ 
			Upper bound	&1 		& $+\infty$ & $+\infty$	& 100 			& 1 		& $+\infty$ \\ \hline
		\end{tabular} }
	\caption{Constraints of the parameter estimation problem}
	\label{tab:Constraints}
	\end{table}

	The parameter estimation problem requires an initial guess as an input to an optimization problem. To ensure that the solution is a global solution multiple initial points have been tested. The initial guesses for the parameter estimation problem are given in Table \ref{tab:InitialGuess}

	\begin{table}[!h]
		\centering
		\adjustbox{max width=\columnwidth}{%
		\begin{tabular}{ lcccccc } 
			\hline
			Parameter		&$k_m$ 	& $D_i$ 	& $D_e^M$ 	& $m_{total}$	& $\tau$ 	& $\sigma$ \\  \hline
			Initial guess	&1	  	& 3 	  	& 1 		& 77 		 	& 0.65 	   	& 0.1 \\  \hline
		\end{tabular} }
		\caption{Constraints of the parameter estimation problem}
		\label{tab:InitialGuess}
	\end{table}

@@

	\begin{figure*}[!h]
		\centering
		\begin{subfigure}[b]{0.8\textwidth}
			\centering
			\includegraphics[trim = 3cm 11cm 2.5cm 2.1cm,clip,width=\textwidth]{/Results_estimation/Fitting_LUKE_T40_P200.pdf}
			\caption{Experiment at $40^\circ C$ and $200$ bar}
		\end{subfigure}
		\hfill
		\begin{subfigure}[b]{0.8\textwidth}
			\centering
			\includegraphics[trim = 3cm 11cm 2.5cm 2.1cm,clip,width=\textwidth]{/Results_estimation/Fitting_LUKE_T50_P200.pdf}
			\caption{Experiment at $50^\circ C$ and $200$ bar}
		\end{subfigure}
		\hfill
		\begin{subfigure}[b]{0.8\textwidth}
			\centering
			\includegraphics[trim = 3cm 11cm 2.5cm 2.1cm,clip,width=\textwidth]{/Results_estimation/Fitting_LUKE_T40_P300_org.pdf}
			\caption{Experiment at $40^\circ C$ and $300$ bar}
		\end{subfigure}
		\hfill
		\begin{subfigure}[b]{0.8\textwidth}
			\centering
			\includegraphics[trim = 3cm 11cm 2.5cm 2.1cm,clip,width=\textwidth]{/Results_estimation/Fitting_LUKE_T50_P300_org.pdf}
			\caption{Experiment at $50^\circ C$ and $300$ bar}
		\end{subfigure}
		\caption{Results of parameter fitting, with estimation of the initial state}
		\label{fig: estimation_results}
	\end{figure*}

@@

	Values of the internal diffusion coefficients are distinguished for each experiment, with an increasing trend with respect to the fluid density. The order of magnitude of $D_i$ obtained from the optimization is similar to values found by other researcher (\todo[fancyline]{REF}).

	\begin{figure*}[!h]
	\centering
	\begin{subfigure}[b]{0.85\textwidth}
		\centering
		\includegraphics[trim = 1cm 2cm 2.5cm 1cm,clip,width=\textwidth]{/Results_estimation/Trend_Lines_order_1.pdf}
		\caption{First order polynomial regression of fitted parameters as a function of fluid density $\rho_f$}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.85\textwidth}
		\centering
		\includegraphics[trim = 1cm 2cm 2.5cm 1cm,clip,width=\textwidth]{/Results_estimation/Trend_Lines_order_2.pdf}
		\caption{Second order polynomial regression of fitted parameters as a function of fluid density $\rho_f$}
	\end{subfigure}
	\caption{Results of parameter fitting, with estimation of the initial state}
	\end{figure*}

>>>>

    The parameter estimation problem was solved by fitting the process model to the dataset given in the Table \ref{tab: Yield_data}. Each time-series was fitted to the model separately. As a result of fitting, the following parameters were obtained:

@@

	To ensure that parameters found by the optimizer do not reach unrealistic values, an additional set of inequality constraints is introduced. The lower and upper bounds for each parameter are given in Table \ref{tab:Constraints}. The same Table contains the initial guesses for the optimizer. To ensure that the solution is a global solution multiple initial points have been tested.

	\begin{table}[!h]
		\centering
		\adjustbox{max width=\columnwidth}{%
		\begin{tabular}{ lcccccc }
			\hline 
			Parameter		&$k_m$ 	& $D_i$ 	& $D_e^M$ 	& $m_{total}$	& $\tau$ 	& $\sigma$ \\  \hline
			Lower bound		&0	  	& 0 	  	& 0 		& 76 		 	& 0 	   	& 0 \\ 
			Upper bound		&1 		& $+\infty$ & $+\infty$	& 100 			& 1 		& $+\infty$ \\ 
			Initial guess	&1	  	& 3 	  	& 1 		& 77 		 	& 0.65 	   	& 0.1 \\  \hline
		\end{tabular} }
	\caption{Constraints of the parameter estimation problem}
	\label{tab:Constraints}
	\end{table}

%	The parameter estimation problem requires an initial guess as an input to an optimization problem. To ensure that the solution is a global solution multiple initial points have been tested. The initial guesses for the parameter estimation problem are given in Table \ref{tab:InitialGuess}

%	\begin{table}[!h]
%		\centering
%		\adjustbox{max width=\columnwidth}{%
%		\begin{tabular}{ lcccccc } 
%			\hline
%			Parameter		&$k_m$ 	& $D_i$ 	& $D_e^M$ 	& $m_{total}$	& $\tau$ 	& $\sigma$ \\  \hline
%			Initial guess	&1	  	& 3 	  	& 1 		& 77 		 	& 0.65 	   	& 0.1 \\  \hline
%		\end{tabular} }
%		\caption{Constraints of the parameter estimation problem}
%		\label{tab:InitialGuess}
%	\end{table}

@@

@@

	Values of the internal diffusion coefficients are distinguished for each experiment, with an increasing trend with respect to the fluid density. The order of magnitude of $D_i$ obtained from the optimization is similar to values found by other researchers. \citet{Reverchon1996} performed the parameter estimation for the extraction process of sage oil from seeds and reported $D_i \approx 6e-13$.

	The axial diffusion coefficient obtained from the optimizer seems to be quite high if compared to the internal diffusion coefficients. Nevertheless, the values of $D_e^M$ have similar order of magnitude as reported in literature, for example \citet{ReisVasco2000}. The relatively high values of the axial diffusion coefficients might be justified by taking into account low values of flow rate used in all the experiments.

	As the total amount of the oil present in the system is unknown, and as such it has been obtained from the optimizer. Given the dataset and the process model, the optimizer found that all most all the oil was extracted out of the seeds and that the best fit is obtained if the total mass of the oil reach a lower bound equal to 76 g.

	As explained above, and in Appendix \ref{CH: IC_BC}, the linear distribution of can be fully determined if the parameter $\tau$ is known. The split ratio for experiments conducted at $40 ^\circ C$ / 300 bar and $50 ^\circ C$ / 300 bar are similar to each other and close to 0.64. The experiments performed at $40 ^\circ C$ / 200 bar and $50 ^\circ C$ / 200 bar have $\tau$ equals to $0.69$ and $0.74$, respectively.

	The noise present in each dataset is quantified by parameter $\sigma$. It can be observed dataset obtained at $40 ^\circ C$ / 200 bar, $50 ^\circ C$ / 200 bar and $40 ^\circ C$ / 300 bar have similar value of $\sigma\approx0.34$. The visual investigation of the dataset obtained at $50 ^\circ C$ / 300 bar allow to explain higher value of $\sigma$ for this dataset. This dataset have higher deviation of the data points from the simulated curve.

		\begin{figure}[!h]
		\centering
		\begin{subfigure}[b]{\columnwidth}
			\centering
			\includegraphics[trim = 3cm 11cm 2.5cm 2.1cm,clip,width=\textwidth]{/Results_estimation/Fitting_LUKE_T40_P200.pdf}
			\caption{Experiment at $40^\circ C$ and $200$ bar}
		\end{subfigure}
		\begin{subfigure}[b]{\columnwidth}
			\centering
			\includegraphics[trim = 3cm 11cm 2.5cm 2.1cm,clip,width=\textwidth]{/Results_estimation/Fitting_LUKE_T50_P200.pdf}
			\caption{Experiment at $50^\circ C$ and $200$ bar}
		\end{subfigure}
		\begin{subfigure}[b]{\columnwidth}
			\centering
			\includegraphics[trim = 3cm 11cm 2.5cm 2.1cm,clip,width=\textwidth]{/Results_estimation/Fitting_LUKE_T40_P300_org.pdf}
			\caption{Experiment at $40^\circ C$ and $300$ bar}
		\end{subfigure}
		\begin{subfigure}[b]{\columnwidth}
			\centering
			\includegraphics[trim = 3cm 11cm 2.5cm 2.1cm,clip,width=\textwidth]{/Results_estimation/Fitting_LUKE_T50_P300_org.pdf}
			\caption{Experiment at $50^\circ C$ and $300$ bar}
		\end{subfigure}
		\caption{Results of parameter fitting, with estimation of the initial state}
		\label{fig: estimation_results}
	\end{figure}

	The results obtained from the parameter estimation are used to develop a generalized process model. This model expresses the parameters $D_i$ and $D_e^M$ as functions of the operating conditions, such as temperature (T), pressure (P), and flow rate (F). Since the experiments were performed under varying temperature and pressure, the influence of these variables is taken into account. To simplify the problem, the correlations are written as functions of fluid density rather than temperature and pressure.

	While the experimental procedure and device can affect $\tau$ and $\sigma$, these parameters are not further considered. They are used only to define the initial conditions and are not included in the process model. Additionally, since the optimizer found that $k_m$ and $m_{total}$ have the same value, they are assumed to be constants and not considered further.

	The other two parameters ($D_i$ and $D_e^M$) are subject to regression to obtain correlations in terms of density. Figure \ref{fig:Regression} shows the first- and second-order polynomials used to fit $D_i$ and $D_e^M$ as a function of fluid density. Interestingly, these two parameters exhibit opposite behavior. The internal diffusion coefficient increases with increasing fluid density, whereas the axial diffusion coefficient decreases.

	\begin{figure}[!h]
	\centering
	\begin{subfigure}[b]{\columnwidth}
		\centering
		\includegraphics[trim = 10.5cm 11cm 2.5cm 1cm,clip,width=\textwidth]{/Results_estimation/Trend_Lines_order_1.pdf}
		\caption{First order polynomial regression of fitted parameters as a function of fluid density $\rho_f$}
	\end{subfigure}
	\begin{subfigure}[b]{\columnwidth}
		\centering
		\includegraphics[trim = 10.5cm 11cm 2.5cm 1cm,clip,width=\textwidth]{/Results_estimation/Trend_Lines_order_2.pdf}
		\caption{Second order polynomial regression of fitted parameters as a function of fluid density $\rho_f$}
	\end{subfigure}
	\caption{Results of parameter fitting, with estimation of the initial state} 
	\label{fig:Regression}
	\end{figure}
